A game theoretic model of the behavioural gaming that takes place at the EMS - ED interface\tnoteref{t1}

### 2022-04-25 reviews/EJOR/01/Response to Reviewers.tex

\documentclass{article}

\title{Response to reviewers}
\author{A game theoretic model of the behavioural gaming that takes \\
place at the ED-EMS interface}

\begin{document}
    \maketitle

    The authors would like to thank all the reviewers for their
    constructive feedback and for
    the time they took to review the manuscript.
    Their suggestions were extremely helpful and we believe this has resulted
    in a much improved manuscript.
    Below is a point by point response and summary of changes made, which have
    been grouped together if reviewers provided similar
    suggestions/improvements.

    \line(1, 0){0.92\textwidth}
    \begin{itemize}
        \item Reviewer 1 - Comment 2: \textit{Combine Sections 2 and 5
        together to demonstrate the underlying queueing models and the
        EMS - ED interface counterpart.
        The current arrangement is not easy to follow for finding the right
        explanations for each part of the assumptions adopted in the
        article.}

        \item Reviewer 3 - Comment 1: \textit{Sections 1 and 2 are
        difficult to read.
        It would be better to focus on the main contribution and its
        motivation in Section 1, and to put the literature review in a
        separate Section.}
    \end{itemize}
    The paper has been completely restructured.
    Sections 2 and 5 have been combined and the game theoretic model is now
    presented along with the results in one section.
    In addition, section 1 contains only the introduction while the
    literature review has moved to section 2.
    Finally, the queueing model section now comes before the game theoretic
    section.
    The current structure is: Introduction; Literature Review; A queueing
    model for the ED-EMS interface; Strategic manipulation of the ED-EMS
    interface; Results.
    We thank the reviewers for their suggestions and hope that the major
    restructuring now ensures that the manuscript flows in a much more coherent
    way.

    \line(1, 0){0.92\textwidth}
    \begin{itemize}
        \item Reviewer 3 - Comment 2: \textit{Since the application in mind
        is the EMS-ED interface, it would be better to write the paper
        immediately from this perspective, instead of first abstracting it
        and then applying it to that problem.}
        \item Reviewer 3 - Comment 8: \textit{Figs 7-11: Row player and
        Column Player are unclear: hospital 1 and hospital 2 would be a
        better choice I presume.}
    \end{itemize}
    The paper is presented with the application of the model as the focus.
    All sections are now presented from the perspective of the EMS-ED
    interface.
    In the game theoretic section instead of row/column player we now use
    hospital A and hospital B.

    \line(1, 0){0.92\textwidth}
    \begin{itemize}
        \item Reviewer 1 - Comment 4: \textit{To highlight the contributions
        of this work, it is necessary to compare the novelty of this work
        with previous results in the literature.}
    \end{itemize}
    In the new literature review section we now directly compare the results
    from previous work with the results from this work to highlight the
    contributions of this work.

    \line(1, 0){0.92\textwidth}
    \begin{itemize}
        \item Reviewer 3 - Comment 6: \textit{Figure 3: about solving the
        Markov chain: the Markov chain is truncated to solve it. Therefore,
        probably N and M are not too small. What about the numerical
        complexity? Because if you just solve it, you have to invert the
        matrix? Is there some structure in the Markov chain that can be
        exploited if N and M would be large?}
        \item Reviewer 3 - Comment 7: \textit{Figures 4-6: why do you
        compare your calculated values with simulation?
        Is it to evaluate the truncations?
        Or are you not sure about the formulas?}
    \end{itemize}
    This observation was really helpful and lead to us conducting more numerical
    experiments to clarify the point of truncating the state space.
    As the values of \(N\) and \(M\) increase, the truncated Markov chain
    becomes more complex to solve.
    We have touched upon building a closed-form formula for the steady state
    probabilities but that is still a work in progress and not something we are
    able to include in this manuscript.
    Figures 4-6 are there to demonstrate the accuracy of our constructed
    model against its discrete event simulation equivalent and these have been
    expanded on.
    An additional paragraph is now included that discusses the truncation of
    the model
    and the previous figures have been replaced with ones that compare the
    results from the truncated Markov model with the results from the
    untruncated simulation model.
    We feel that these extra numeric results
    help clarify the paper and are thankful for the reviewer comments.

    \line(1, 0){0.92\textwidth}
    \begin{itemize}
        \item Reviewer 3 - Comment 3: \textit{In section 2, a diagram
        of the full system (including the two queues and the EMS) would
        increase readability.}
    \end{itemize}
    A full system diagram is now added in the game theoretic model section.

    \line(1, 0){0.92\textwidth}
    \begin{itemize}
        \item Reviewer 3 - Comment 9: \textit{Page 22 scenario at bottom of
        the page: this scenario does not seem like a realistic scenario:
        the mean number of patients offered per time unit is larger than
        that can be treated, meaning that many patients would be rejected
        by the hospital. Please comment on the reality of the selected
        scenarios}
    \end{itemize}
    The queueing systems in the particular example are meant to be extremely
    busy systems.
    In the new manuscript this is more clearly stated.

    \line(1, 0){0.92\textwidth}
    \begin{itemize}
        \item Reviewer 3 - Comment 5: \textit{A key modelling assumption is
        that a fraction \(p_A\) of patients is distributed to hospital A.
        Some comments on how realistic this is should be added. Do
        distributers have a choice? Don't they just have to choose the
        nearest hospital? Do they really do it randomly, don' t they take
        into account queues at the hospitals (e.g. join the shortest
        ambulance queue)? what about urgency of patients? I would expect
        patients that come in are prioritized since they are probably more
        urgent.}
    \end{itemize}
    An additional paragraph has been added that touches upon the possibility of
    more complex queueing models where the hospital distance or patient's
    priority type are considered.

    \line(1, 0){0.92\textwidth}
    \begin{itemize}
        \item Reviewer 3 - Comment 4: \textit{The authors claim their model
        has applications in a range of sectors and settings, but have given
        no examples.}
    \end{itemize}
    Examples of other applications have been added in the conclusion section
    of the manuscript.

    \line(1, 0){0.92\textwidth}
    \begin{itemize}
        \item Reviewer 1 - Comment 3: \textit{In the game-theoretic
        formulation and analysis part, it is better to present the
        essential components of a normal-form representation, including the
        players in the game, the strategies available to each player and the
        payoff received by each player for each combination of strategies
        that could be chosen by the players, etc.}
    \end{itemize}
    In the new manuscript the normal form game is explicitly defined in the
    game theoretic section. The players, strategies and payoffs are clearly
    defined.

    \line(1, 0){0.92\textwidth}
    \begin{itemize}
        \item Reviewer 1 - Comment 1: \textit{Rewrite the Highlights (for
        review) of the paper, and really capture the sense of work.}
    \end{itemize}
    Highlights have been rewritten to be more explicit in terms of the work
    that was done.

    \line(1, 0){0.92\textwidth}
    \begin{itemize}
        \item Reviewer 1 - Comment 5: \textit{There are some mistakes/typos
        with English grammar and writing in the article.}
    \end{itemize}
    We have thoroughly checked the grammar and spelling of the manuscript.

\end{document}

### 2022-04-25 sections/model_overview/main.tex
Address all reviewers comments (#31)
* Prepare for addressing reviewers comments (#18)

* Add comments of reviewers

`main.md` consists of the comments of all three reviewers and the editor

* Include editor's comments

* Change vocabulary to be more focused on the application (#33)

* Change vocabulary in game section

* Minor typo

Fix proselint failure by removing the word very from a comment

* Change vocabulary in methodology section

* Additional changes in game section

- Replace waiting zone 1 with hospital
- Replace waiting zone 2 with parking space
- Use type 2 patients everywhere in

The strategies of the two EDs are the range of thresholds that they can choose
from and their utilities is the proportion of individuals whose time in the 
system is within a predetermined target time.
The EMS has to decide how to distribute its patients among the two EDs so that 
the weighted combination of the ambulance blocking time and the percentage of 
lost ambulances is minimised. 
This can be illustrated by figure \ref{fig:diagram_of_game_theoretic_model}.
The interaction between the two
EDs is a normal form game that is then used to inform the decision of the EMS.
Note that the formulated game here assumes that prior to making a choice the 
EMS knows the strategies that each ED is playing (figure 
\ref{fig:imperfect_info_game}). 
This corresponds to reacting to experienced delays.

@@

Both hospitals follow a queueing model with two waiting spaces for 
individuals. 
The first waiting space (i.e. the waiting space of the hospital) is where the 
patients queue right before receiving 
their service and has a queue capacity of \( N - C \), where \(N\) is the total 
capacity of the hospital and \(C\) is the number of healthcare 
professionals able to see them.
The second waiting space (i.e. the parking space for ambulances) is where 
ambulances, that are sent from the
ambulance service, stay until their patients are allowed to enter the hospital.
The parking space has a capacity of \(M\) and no servers.
This is shown diagrammatically in Figure~\ref{fig:diagram_of_queueing_system}.

@@

Here, \(\alpha\) represents the ``importance'' of each objective,
where high \(\alpha\) indicates a higher weight on the proportion of lost 
individuals and smaller \(\alpha\) a higher weight on the time blocked. 
The choice of \(p_A\) and \(p_B\) rely solely on equation
\ref{eq:obj_distributor}.
Note that the current system is modelled using unobservable queues which is not
necessarily an unrealistic approach~\cite{unobservablequeue}.
Another approach would be to allow the ambulance service to observe the state
of each hospital before deciding which one to join for a given individual.
There are several other factors that can affect the routing of the
patients that are outside the scope of this paper.
For example the distance from each hospital or even the priority level of each
patient may be a significant factor that affects the ambulance service's
decision.

Using either equation (\ref{eq:obj_distributor}) or 
(\ref{eq:obj_queueing_systems}) gives an imperfect information extensive form 
game. 
An imperfect information game is defined as an extensive form game where some 
of the information about the game state is hidden for at least one of the 
players~\cite{Berwanger2008}. In this study the state of the problem that is
hidden is the threshold that each of the hospitals chooses to play.
In other words, each hospital chooses to play a strategy without 
knowing the other hospital's strategy.
The ambulance service then, fully aware of the chosen threshold strategies, 
distributes individuals among the two systems in order to minimise the time 
that its ambulances will be blocked. 
Figure \ref{fig:imperfect_info_game} illustrates this.

>>>>

The strategies of the two EDs are the range of thresholds that they can choose
from and their utilities are the proportions of individuals whose time in the 
system is within a predetermined target time.
The EMS has to decide how to distribute its patients among the two EDs so that 
the weighted combination of the ambulance blocking time and the percentage of 
lost ambulances is minimised. 
This can be illustrated by figure \ref{fig:diagram_of_game_theoretic_model}.
The interaction between the two
EDs is a normal form game that is then used to inform the decision of the EMS.
Note that the formulated game here assumes that prior to making a choice the 
EMS knows the strategies that each ED is playing (figure 
\ref{fig:imperfect_info_game}). 
This corresponds to reacting to experienced delays.

@@

Both hospitals follow a queueing model with two waiting spaces for 
individuals. 
The first waiting space (i.e. the waiting space of the hospital) is where the 
patients queue right before receiving 
their service and has a queue capacity of \( N - C \), where \(N\) is the total 
capacity of the hospital and \(C\) is the number of healthcare 
professionals able to see them.
The second waiting space (i.e. the parking space for ambulances) is where 
ambulances, that are sent from the
EMS distributor, stay until their patients are allowed to enter the hospital.
The parking space has a capacity of \(M\) and no servers.
This is shown diagrammatically in Figure~\ref{fig:diagram_of_queueing_system}.

@@

Here, \(\alpha\) represents the ``importance'' of each objective,
where a high \(\alpha\) indicates a higher weight on the proportion of lost 
individuals and a smaller \(\alpha\) a higher weight on the time blocked. 
The choice of \(p_A\) and \(p_B\) rely solely on equation
\ref{eq:obj_distributor}.
Note that the current system is modelled using unobservable queues which is not
necessarily an unrealistic approach~\cite{unobservablequeue}.
Another approach would be to allow the ambulance service to observe the state
of each hospital before deciding which one to join for a given individual.
There are several other factors that can affect the routing of the
patients that are outside the scope of this paper.
For example the distance from each hospital or even the priority level of each
patient may be a significant factor that affects the ambulance service's
decision.

Using either equation (\ref{eq:obj_distributor}) or 
(\ref{eq:obj_queueing_systems}) gives an imperfect information extensive form 
game. 
An imperfect information game is defined as an extensive form game where some 
of the information about the game state is hidden for at least one of the 
players~\cite{Berwanger2008}. In this study the state of the problem that is
hidden is the threshold that each hospital ED chooses to play.
In other words, each hospital chooses to play a strategy without 
knowing the other hospital's strategy.
The ambulance service then, fully aware of the chosen threshold strategies, 
distributes individuals among the two systems in order to minimise the time 
that its ambulances will be blocked. 
Figure \ref{fig:imperfect_info_game} illustrates this.

### 2022-04-25 sections/queueing_model/blocking_time/main.tex
Address several TODOs (#41)
* Update highlights

* Address several todos

* Add a final TODO to be fixed

* Add example of another application of the model

* Fix a typo

* Add script and notebook of previous plots

* Rename jupyter notebook

* Change plots axis labels

* Scale payoff matrices to be more presentable

* Remove unnecessary equations

* Minor additional changes

* Address the high traffic intensity
Address all reviewers comments (#31)
* Prepare for addressing reviewers comments (#18)

* Add comments of reviewers

`main.md` consists of the comments of all three reviewers and the e

% TODO: Possibly replace the contents of this section with the ones for the 
% closed form formula of the blocking time
% Currently the direct approach one is shown

Blocking time is the amount of time that type 2 patients wait in the parking 
space before they are allowed to proceed to the hospital.
Unlike the waiting time, the blocking time is only calculated for type 2 
individuals.  
That is because type 1 individuals cannot be blocked. 
Thus, one only needs to consider the pathway of type 2 individuals to get the 
mean blocking time of the system. 
The mean blocking time can by calculated using:

@@

Figure \ref{fig:markov_vs_des_blocking_time_comparison} illustrates a comparison 
between the formulas that arise from the Markov chain model and the equivalent 
values of the blocking time extracted from discrete event simulation
(appendix~\ref{sec:appendix_des}).
The blocking time is calculated using both methods for a range of values of
\(\lambda_2\).
The figure is used to demonstrate the accuracy of the blocking time formula of
the constructed queueing model as well as the effect of truncating the model.
The simulation was ran 100 times and the recorded mean blocking time at each 
iteration is used to populate the violin plots.
Similar to figure \ref{fig:markov_vs_des_waiting_time_comparison_overall}, these
plots shows a comparison between the calculated mean blocking time using Markov 
chain, using a truncated simulation and using a simulation without the
artificial parameters \(N\) and \(M\).
The blocking times generated by the truncated simulation match the ones
generated by the Markov chains model.
Note that this comparison includes only type 2 individuals since type 1 
individuals cannot be blocked.

>>>>

Blocking time is the amount of time that type 2 patients wait in the parking 
space before they are allowed to proceed into the hospital.
Unlike the waiting time, the blocking time is only calculated for type 2 
individuals.  
That is because type 1 individuals cannot be blocked. 
Thus, one only needs to consider the pathway of type 2 individuals to get the 
mean blocking time of the system. 
The mean blocking time can by calculated using:

@@

Figure \ref{fig:markov_vs_des_blocking_time_comparison} illustrates a comparison 
between the formulas that arise from the Markov chain model and the equivalent 
values of the blocking time extracted from discrete event simulation
(appendix~\ref{sec:appendix_des}).
The blocking time is calculated using both methods for a range of values of
\(\lambda_2\).
The figure is used to demonstrate the accuracy of the blocking time formula of
the constructed queueing model as well as the effect of truncating the model.
The simulation was run 100 times and the recorded mean blocking time at each 
iteration is used to populate the violin plots.
Similar to figure \ref{fig:markov_vs_des_waiting_time_comparison_overall}, these
plots shows a comparison between the calculated mean blocking time using Markov 
chain, using a truncated simulation and using a simulation without the
artificial parameters \(N\) and \(M\).
The blocking times generated by the truncated simulation match the ones
generated by the Markov chains model.
Note that this comparison includes only type 2 individuals since type 1 
individuals cannot be blocked.

### 2022-04-25 sections/queueing_model/main.tex
Investigate the truncation effect (#34)
* Get initial plots of truncation effects

Plots added on unbounded simulation VS Markov chain (while
increasing capacity):
- Waiting time
- Blocking time
- Proportion of individuals within time

* Add bounded simulation to same plot

* Write script to generate duration for each method

* Get timing data when M=50

Times for N = (10, 30, 50, inf) for markov (waiting, blocking and
proportion formula) and simulation (1 trial, 20 trials)

* Get timing data when M=5

Times for N = (10, 30, 50, inf) for markov (waiting, blocking and
proportion formula) and si

\section{A queueing model for the ED-EMS interface}\label{sec:queueing_model}
In this section, a more in-depth explanation of the queueing model shown in 
figure \ref{fig:diagram_of_queueing_system} will be given.
This is a queueing model that consists of two waiting zones; the parking space
and the hospital waiting space.

@@

The model consists of two types of individuals; type 1 and type 2.
Type 2 individuals are patients arriving in ambulances who can be blocked 
(usually patients that are deemed not to be critical) 
and type 1 individuals are individuals arriving from other sources 
(e.g walk-in patients, urgent patients from either walk-ins or ambulances).
Type 1 individuals arrive instantly at the hospital's waiting space and wait to
receive their service. 
Type 2 individuals arrive at the parking space and wait there until they are 
allowed to move to the hospital.
They are allowed to proceed only when the number of 
patients in the hospital is less than the
pre-determined threshold \(T\).
When the number of individuals is equal to or exceeds this threshold, all 
type 2 patients that arrive will stay 
\textit{blocked} in the parking space until the number of patients in 
the hospital falls below \(T\). 
This is shown diagrammatically in Figure \ref{fig:diagram_of_queueing_system}.
The parameters of the described queueing model are:

@@

>>>>

\section{A queueing model for the ED-EMS interface}\label{sec:queueing_model}
In this section, a more in-depth explanation of the queueing model shown in 
figure \ref{fig:diagram_of_queueing_system} will be given.
This is a queueing model that consists of two waiting zones: the parking space
and the hospital waiting space.

@@

The model consists of two types of individuals; type 1 and type 2.
Type 2 individuals are patients arriving in ambulances who can be blocked 
(usually patients that are deemed not to be critical) 
and type 1 individuals are individuals arriving from other sources 
(e.g walk-in patients, urgent patients from either walk-ins or ambulances).
Type 1 individuals arrive instantly at the hospital's waiting space and wait to
receive their service. 
Type 2 individuals arrive at the parking space and wait there until they are 
allowed to move into the hospital.
They are allowed to proceed only when the number of 
patients in the hospital is less than the
pre-determined threshold \(T\).
When the number of individuals is equal to or exceeds this threshold, all 
type 2 patients that arrive will stay 
\textit{blocked} in the parking space until the number of patients in 
the hospital falls below \(T\). 
This is shown diagrammatically in Figure \ref{fig:diagram_of_queueing_system}.
The parameters of the described queueing model are:

@@

\input{sections/queueing_model/truncation_effect/main.tex}
